---
title: "Larval density dependence"
author: "Andreas Kudom, Dave Hemprich-Bennett, Talya Hackett"
date: "06/07/2022"
output: bookdown::html_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(here)

# Rerun scripts
source(here('scripts', '1_larval_growth_rates.R'))
source(here('scripts', '2_larval_mortality.R'))       

# read in formatted df
larval_density_df <- read_csv(here('data', 'processed_data', '1_density_and_growth.csv')) 
sp <- larval_density_df %>% pull(species) %>% unique()
```

# Introduction

Andreas ran an experiment where containers had `r larval_density_df %>% pull(larval_density) %>% unique()` *Anopheles gambiae* eggs placed within them. 
`r larval_density_df %>% pull(replicate_no) %>% unique() %>% (length)` simultaneous replicates of different larval densities were performed, over `r larval_density_df %>% pull(date_of_experiment) %>% unique() %>% length()` iterations. 
Iterations either had `r sp[1]` or `r sp[2]` varieties used.  
Three sets of variables were measured:  

1. Number of larvae per surviving until pupation per replicate
2. For larvae that survived until pupation, number of days between hatching and:
    + pupation
    + eclosion
3. Wing length for a subset of larvae that survived until eclosion.


# Survival rates

Increasing the density of larvae in a container decreased the proportion of those larvae who survived (Figure \@ref(fig:prop-surv)), although the highest number of larvae survived in the containers with 60 larvae (Figure \@ref(fig:n-surv)).

```{r prop-surv, fig.cap="Proportion of larvae surviving."}
proportion_survival_boxplot
```



```{r n-surv, fig.cap="Number of larvae surviving.", echo = F}
number_surviving_boxplot
```


